\textbf{The status of the archaeal and bacterial census: an update}

### 2016-03-14 submission/Schloss_Census2_mBio_2016.tex
Fresh remake on axiom

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={The status of the archaeal and bacterial census: an update},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

\textbf{\emph{Focusing the census by environment.}} We were able to
assign 89.3 and 95.1\% of the sequences to one of seven broad
environmental categories based on the metadata that accompanied the
SILVA database (Tables 1). Across these broad categories there was wide
variation in the number of sequences that have been sampled. Among
bacterial sequences, the three best represented groups were from
zoonotic (N=804,585), aquatic (N=214,085), and built environment
(N=108,799) sources. Among the archaeal sequences the three best
represented groups were the same, but ordered differently: aquatic
(N=34,400), built environment (N=7,286), and zoonotic (N=5,597) (Figure
1C,D)). For both domains, soil samples were the fourth most represented
category (bacteria: 74,870; archaea: 2,517). The orders of these
categories was surprising considering soil and aquatic environments
harbor the most microbial biomass and biodiversity (21). In spite of
wide variation in sequencing depth and coverage (Table 1), the
interquartile range across the fine-level categories for the bacterial
OTU coverage only varied between 34.5 to 40.0 (median coverage=36.7\%).
The interquartile range in the OTU coverage by environment for the
archaeal data was 41.5 to 53.1 (median coverage=44.9\%). The archaeal
coverage was higher than that of the bacterial OTU coverage for all
categories except the food-associated, plant surface, and other
invertebrate categories. Across all categories, the bacterial and
archaeal sequencing data represented a limited number of phyla (Figure
4). Among the bacterial data, the fine-scale categories were dominated
by Proteobacteria (N=24), Firmicutes (N=2), and Actinobacteria (N=1) and
among the archaeal data, they were dominated by Euryarchaeota (N=16),
Thaumarchaeota (N=10), and Aenigmarchaeota (N=1). Regardless, there were
clear phylum-level signatures that differentiated the various
categories. Within each of the bacterial and archaeal phyla, there was
considerable variation in the relative abundance of each across the
categories confirming that taxonomic signatures exist to differentiate
different environments even at a broad taxonomic level.

@@

Novel technologies such as single-cell genomics, metagenomics, and
algorithms to recover full-length sequences from new sequencing
platforms have demonstrated promise in circumventing previous
limitations in identifying new OTUs. Using EMIRGE to assemble fragmented
16S rRNA gene amplicons may allow us to obtain deep coverage of
communities; however, it is still unclear how faithful the assembled
sequence is to that of the original organism. Additional sequencing
technologies also offer the ability to directly generate full-length
sequences, such as PacBio and potentially Oxford Nanopore. Initial
application of PacBio to sequencing full-length fragments suggests that
the sequences suffered from a high error rate (39). To obtain a more
direct investigation of rare organisms, microbiologists are developing
novel cultivation and single cell genomics techniques (\textbf{???},
40--42). The ability to enrich or select for specific populations using
these approaches could limit the need for redundant brute force
sequencing. These approaches are still in active development, and we
hope that through continuous refinement, they may allow us to
significantly improve the coverage of OTUs in public databases.

@@

\textbf{\emph{Data analysis.}} Our analysis made use of ARB (OS X v.6.0)
(43), mothur (v.1.37.0) (44), and R (v.3.2.3) (47). Within R we utilized
the knitr (v.1.12.3), wesanderson (v.0.3.3.99), and openxlsx (v. 2.4.0)
packages. A reproducible version of this manuscript including data
extraction and processing is available at
\url{https://www.github.com/SchlossLab/Schloss_Census2_mBio_2016}.

@@

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\hypertarget{ref-McGill2007}{}
1. \textbf{McGill BJ}, \textbf{Etienne RS}, \textbf{Gray JS},
\textbf{Alonso D}, \textbf{Anderson MJ}, \textbf{Benecha HK},
\textbf{Dornelas M}, \textbf{Enquist BJ}, \textbf{Green JL}, \textbf{He
F}, \textbf{Hurlbert AH}, \textbf{Magurran AE}, \textbf{Marquet PA},
\textbf{Maurer BA}, \textbf{Ostling A}, \textbf{Soykan CU},
\textbf{Ugland KI}, \textbf{White EP}. 2007. Species abundance
distributions: Moving beyond single prediction theories to integration
within an ecological framework. Ecology Letters \textbf{10}:995--1015.
doi:\href{https://doi.org/10.1111/j.1461-0248.2007.01094.x}{10.1111/j.1461-0248.2007.01094.x}.

\hypertarget{ref-Hubbell2001}{}
2. \textbf{Hubbell SP}. 2001. A Unified Theory of Biodiversity and
Biogeography. Princeton University Press, Princeton.

\hypertarget{ref-Konstantinidis2006}{}
3. \textbf{Konstantinidis KT}, \textbf{Ramette A}, \textbf{Tiedje JM}.
2006. The bacterial species definition in the genomic era. Philosophical
Transactions of the Royal Society B: Biological Sciences
\textbf{361}:1929--1940.
doi:\href{https://doi.org/10.1098/rstb.2006.1920}{10.1098/rstb.2006.1920}.

\hypertarget{ref-Oren2013}{}
4. \textbf{Oren A}, \textbf{Garrity GM}. 2013. Then and now: A
systematic review of the systematics of prokaryotes in the last
80~years. Antonie van Leeuwenhoek \textbf{106}:43--56.
doi:\href{https://doi.org/10.1007/s10482-013-0084-1}{10.1007/s10482-013-0084-1}.

\hypertarget{ref-Brosius1978}{}
5. \textbf{Brosius J}, \textbf{Palmer ML}, \textbf{Kennedy PJ},
\textbf{Noller HF}. 1978. Complete nucleotide sequence of a 16S
ribosomal RNA gene from \emph{Escherichia coli.} Proceedings of the
National Academy of Sciences \textbf{75}:4801--4805.
doi:\href{https://doi.org/10.1073/pnas.75.10.4801}{10.1073/pnas.75.10.4801}.

\hypertarget{ref-Schloss2004}{}
6. \textbf{Schloss PD}, \textbf{Handelsman J}. 2004. Status of the
microbial census. Microbiology and Molecular Biology Reviews
\textbf{68}:686--691.
doi:\href{https://doi.org/10.1128/mmbr.68.4.686-691.2004}{10.1128/mmbr.68.4.686-691.2004}.

\hypertarget{ref-Dykhuizen1998}{}
7. \textbf{Dykhuizen DE}. 1998. Santa Rosalia revisited: Why are there
so many species of bacteria? Antonie van Leeuwenhoek \textbf{73}:25--33.
doi:\href{https://doi.org/10.1023/a:1000665216662}{10.1023/a:1000665216662}.

\hypertarget{ref-Curtis2002}{}
8. \textbf{Curtis TP}, \textbf{Sloan WT}, \textbf{Scannell JW}. 2002.
Estimating prokaryotic diversity and its limits. Proceedings of the
National Academy of Sciences \textbf{99}:10494--10499.
doi:\href{https://doi.org/10.1073/pnas.142680199}{10.1073/pnas.142680199}.

\hypertarget{ref-HMP2012}{}
9. \textbf{The Human Microbiome Consortium}. 2012. Structure, function
and diversity of the healthy human microbiome. Nature
\textbf{486}:207--214.
doi:\href{https://doi.org/10.1038/nature11234}{10.1038/nature11234}.

\hypertarget{ref-Gilbert2014}{}
10. \textbf{Gilbert JA}, \textbf{Jansson JK}, \textbf{Knight R}. 2014.
The Earth Microbiome Project: Successes and aspirations. BMC Biology
\textbf{12}:69.
doi:\href{https://doi.org/10.1186/s12915-014-0069-1}{10.1186/s12915-014-0069-1}.

\hypertarget{ref-AmaralZettler2010}{}
11. \textbf{Amaral-Zettler L}, \textbf{Artigas LF}, \textbf{Baross J},
\textbf{P.A. LB}, \textbf{Boetius A}, \textbf{Chandramohan D},
\textbf{Herndl G}, \textbf{Kogure K}, \textbf{Neal P},
\textbf{Pedrós-Alió C}, \textbf{Ramette A}, \textbf{Schouten S},
\textbf{Stal L}, \textbf{Thessen A}, \textbf{Leeuw J de}, \textbf{Sogin
M}. 2010. A global census of marine microbes, pp. 221--245. \emph{In}
Life in the worlds oceans. Wiley-Blackwell.

\hypertarget{ref-Youssef2009}{}
12. \textbf{Youssef N}, \textbf{Sheik CS}, \textbf{Krumholz LR},
\textbf{Najar FZ}, \textbf{Roe BA}, \textbf{Elshahed MS}. 2009.
Comparison of species richness estimates obtained using nearly complete
fragments and simulated pyrosequencing-generated fragments in 16S rRNA
gene-based environmental surveys. Applied and Environmental Microbiology
\textbf{75}:5227--5236.
doi:\href{https://doi.org/10.1128/aem.00592-09}{10.1128/aem.00592-09}.

\hypertarget{ref-Schloss2010}{}
13. \textbf{Schloss PD}. 2010. The effects of alignment quality,
distance calculation method, sequence filtering, and region on the
analysis of 16S rRNA gene-based studies. PLoS Comput Biol
\textbf{6}:e1000844.
doi:\href{https://doi.org/10.1371/journal.pcbi.1000844}{10.1371/journal.pcbi.1000844}.

\hypertarget{ref-Alivisatos2015}{}
14. \textbf{Alivisatos AP}, \textbf{Blaser MJ}, \textbf{Brodie EL},
\textbf{Chun M}, \textbf{Dangl JL}, \textbf{Donohue TJ},
\textbf{Dorrestein PC}, \textbf{Gilbert JA}, \textbf{Green JL},
\textbf{Jansson JK}, \textbf{Knight R}, \textbf{Maxon ME},
\textbf{McFall-Ngai MJ}, \textbf{Miller JF}, \textbf{Pollard KS},
\textbf{Ruby EG}, \textbf{Taha SA}. 2015. A unified initiative to
harness Earth's microbiomes. Science \textbf{350}:507--508.
doi:\href{https://doi.org/10.1126/science.aac8480}{10.1126/science.aac8480}.

\hypertarget{ref-Liux5f2012}{}
15. \textbf{Li E}, \textbf{Hamm CM}, \textbf{Gulati AS}, \textbf{Sartor
RB}, \textbf{Chen H}, \textbf{Wu X}, \textbf{Zhang T}, \textbf{Rohlf
FJ}, \textbf{Zhu W}, \textbf{Gu C}, \textbf{Robertson CE}, \textbf{Pace
NR}, \textbf{Boedeker EC}, \textbf{Harpaz N}, \textbf{Yuan J},
\textbf{Weinstock GM}, \textbf{Sodergren E}, \textbf{Frank DN}. 2012.
Inflammatory bowel diseases phenotype, textitC. difficile and NOD2
genotype are associated with shifts in human ileum associated microbial
composition. PLoS ONE \textbf{7}:e26284.
doi:\href{https://doi.org/10.1371/journal.pone.0026284}{10.1371/journal.pone.0026284}.

\hypertarget{ref-Kongux5f2012}{}
16. \textbf{Kong HH}, \textbf{Oh J}, \textbf{Deming C}, \textbf{Conlan
S}, \textbf{Grice EA}, \textbf{Beatson MA}, \textbf{Nomicos E},
\textbf{Polley EC}, \textbf{Komarow HD}, \textbf{Murray PR},
\textbf{Turner ML}, \textbf{Segre JA}. 2012. Temporal shifts in the skin
microbiome associated with disease flares and treatment in children with
atopic dermatitis. Genome Research \textbf{22}:850--859.
doi:\href{https://doi.org/10.1101/gr.131029.111}{10.1101/gr.131029.111}.

\hypertarget{ref-Griceux5f2009}{}
17. \textbf{Grice EA}, \textbf{Kong HH}, \textbf{Conlan S},
\textbf{Deming CB}, \textbf{Davis J}, \textbf{Young AC},
\textbf{Bouffard GG}, \textbf{Blakesley RW}, \textbf{Murray PR},
\textbf{Green ED}, \textbf{Turner ML}, \textbf{Segre JA}. 2009.
Topographical and temporal diversity of the human skin microbiome.
Science \textbf{324}:1190--1192.
doi:\href{https://doi.org/10.1126/science.1171700}{10.1126/science.1171700}.

\hypertarget{ref-Griceux5f2010}{}
18. \textbf{Grice EA}, \textbf{Snitkin ES}, \textbf{Yockey LJ},
\textbf{Bermudez DM}, \textbf{Liechty KW}, \textbf{Segre JA},
\textbf{Mullikin J}, \textbf{Blakesley R}, \textbf{Young A}, \textbf{Chu
G}, \textbf{Ramsahoye C}, \textbf{Lovett S}, \textbf{Han J},
\textbf{Legaspi R}, \textbf{Fuksenko T}, \textbf{Reddix-Dugue N},
\textbf{Sison C}, \textbf{Gregory M}, \textbf{Montemayor C},
\textbf{Gestole M}, \textbf{Hargrove A}, \textbf{Johnson T},
\textbf{Myrick J}, \textbf{Riebow N}, \textbf{Schmidt B},
\textbf{Novotny B}, \textbf{Gupti J}, \textbf{Benjamin B},
\textbf{Brooks S}, \textbf{Coleman H}, \textbf{Ho S-l},
\textbf{Schandler K}, \textbf{Smith L}, \textbf{Stantripop M},
\textbf{Maduro Q}, \textbf{Bouffard G}, \textbf{Dekhtyar M},
\textbf{Guan X}, \textbf{Masiello C}, \textbf{Maskeri B},
\textbf{McDowell J}, \textbf{Park M}, \textbf{Thomas PJ}. 2010.
Longitudinal shift in diabetic wound microbiota correlates with
prolonged skin defense response. Proceedings of the National Academy of
Sciences \textbf{107}:14799--14804.
doi:\href{https://doi.org/10.1073/pnas.1004204107}{10.1073/pnas.1004204107}.

\hypertarget{ref-Kirkux5fHarrisux5f2012}{}
19. \textbf{Harris JK}, \textbf{Caporaso JG}, \textbf{Walker JJ},
\textbf{Spear JR}, \textbf{Gold NJ}, \textbf{Robertson CE},
\textbf{Hugenholtz P}, \textbf{Goodrich J}, \textbf{McDonald D},
\textbf{Knights D}, \textbf{Marshall P}, \textbf{Tufo H}, \textbf{Knight
R}, \textbf{Pace NR}. 2012. Phylogenetic stratigraphy in the guerrero
negro hypersaline microbial mat. The ISME Journal \textbf{7}:50--60.
doi:\href{https://doi.org/10.1038/ismej.2012.79}{10.1038/ismej.2012.79}.

\hypertarget{ref-Sogin2006}{}
20. \textbf{Sogin ML}, \textbf{Morrison HG}, \textbf{Huber JA},
\textbf{Welch DM}, \textbf{Huse SM}, \textbf{Neal PR}, \textbf{Arrieta
JM}, \textbf{Herndl GJ}. 2006. Microbial diversity in the deep sea and
the underexplored ``rare biosphere''. Proceedings of the National
Academy of Sciences \textbf{103}:12115--12120.
doi:\href{https://doi.org/10.1073/pnas.0605127103}{10.1073/pnas.0605127103}.

\hypertarget{ref-Whitman1998}{}
21. \textbf{Whitman WB}, \textbf{Coleman DC}, \textbf{Wiebe WJ}. 1998.
Prokaryotes: The unseen majority. Proceedings of the National Academy of
Sciences \textbf{95}:6578--6583.
doi:\href{https://doi.org/10.1073/pnas.95.12.6578}{10.1073/pnas.95.12.6578}.

>>>>

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={The status of the archaeal and bacterial census: an update},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

\textbf{\emph{Focusing the census by environment.}} We were able to
assign 89.3 and 95.1\% of the sequences to one of seven broad
environmental categories based on the metadata that accompanied the
SILVA database (Tables 1). Across these broad categories there was wide
variation in the number of sequences that have been sampled. Among
bacterial sequences, the three best represented groups were from
host-associated (N=804,585), aquatic (N=214,085), and built environment
(N=108,799) sources. Among the archaeal sequences the three best
represented groups were the same, but ordered differently: aquatic
(N=34,400), built environment (N=7,286), and host-associated (N=5,597)
(Figure 1C,D)). For both domains, soil samples were the fourth most
represented category (bacteria: 74,870; archaea: 2,517). The orders of
these categories was surprising considering soil and aquatic
environments harbor the most microbial biomass and biodiversity (21). In
spite of wide variation in sequencing depth and coverage (Table 1), the
interquartile range across the fine-level categories for the bacterial
OTU coverage only varied between 34.5 to 40.0 (median coverage=36.7\%).
The interquartile range in the OTU coverage by environment for the
archaeal data was 41.5 to 53.1 (median coverage=44.9\%). The archaeal
coverage was higher than that of the bacterial OTU coverage for all
categories except the food-associated, plant surface, and other
invertebrate categories. Across all categories, the bacterial and
archaeal sequencing data represented a limited number of phyla (Figure
4). Among the bacterial data, the fine-scale categories were dominated
by Proteobacteria (N=24), Firmicutes (N=2), and Actinobacteria (N=1) and
among the archaeal data, they were dominated by Euryarchaeota (N=16),
Thaumarchaeota (N=10), and Aenigmarchaeota (N=1). Regardless, there were
clear phylum-level signatures that differentiated the various
categories. Within each of the bacterial and archaeal phyla, there was
considerable variation in the relative abundance of each across the
categories confirming that taxonomic signatures exist to differentiate
different environments even at a broad taxonomic level.

@@

Novel technologies such as single-cell genomics, metagenomics, and
algorithms to recover full-length sequences from new sequencing
platforms have demonstrated promise in circumventing previous
limitations in identifying new OTUs. Using EMIRGE to assemble fragmented
16S rRNA gene amplicons may allow us to obtain deep coverage of
communities; however, it is still unclear how faithful the assembled
sequence is to that of the original organism. Additional sequencing
technologies also offer the ability to directly generate full-length
sequences, such as PacBio and potentially Oxford Nanopore. Initial
application of PacBio to sequencing full-length fragments suggests that
the sequences suffered from a high error rate (39). To obtain a more
direct investigation of rare organisms, microbiologists are developing
novel cultivation and single cell genomics techniques ({\textbf{???}},
40--42). The ability to enrich or select for specific populations using
these approaches could limit the need for redundant brute force
sequencing. These approaches are still in active development, and we
hope that through continuous refinement, they may allow us to
significantly improve the coverage of OTUs in public databases.

@@

\textbf{\emph{Data analysis.}} Our analysis made use of ARB (OS X v.6.0)
(43), mothur (v.1.37.0) (44), and R (v.3.2.0) (47). Within R we utilized
the knitr (v.1.10.5), wesanderson (v.0.3.2), and openxlsx (v. 2.4.0)
packages. A reproducible version of this manuscript including data
extraction and processing is available at
\url{https://www.github.com/SchlossLab/Schloss_Census2_mBio_2016}.

@@

\hyperdef{}{references}{\label{references}}
\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\hyperdef{}{ref-McGill2007}{\label{ref-McGill2007}}
1. \textbf{McGill BJ}, \textbf{Etienne RS}, \textbf{Gray JS},
\textbf{Alonso D}, \textbf{Anderson MJ}, \textbf{Benecha HK},
\textbf{Dornelas M}, \textbf{Enquist BJ}, \textbf{Green JL}, \textbf{He
F}, \textbf{Hurlbert AH}, \textbf{Magurran AE}, \textbf{Marquet PA},
\textbf{Maurer BA}, \textbf{Ostling A}, \textbf{Soykan CU},
\textbf{Ugland KI}, \textbf{White EP}. 2007. Species abundance
distributions: Moving beyond single prediction theories to integration
within an ecological framework. Ecology Letters \textbf{10}:995--1015.
doi:\url{http://doi.org/10.1111/j.1461-0248.2007.01094.x}.

\hyperdef{}{ref-Hubbell2001}{\label{ref-Hubbell2001}}
2. \textbf{Hubbell SP}. 2001. A Unified Theory of Biodiversity and
Biogeography. Princeton University Press, Princeton.

\hyperdef{}{ref-Konstantinidis2006}{\label{ref-Konstantinidis2006}}
3. \textbf{Konstantinidis KT}, \textbf{Ramette A}, \textbf{Tiedje JM}.
2006. The bacterial species definition in the genomic era. Philosophical
Transactions of the Royal Society B: Biological Sciences
\textbf{361}:1929--1940.
doi:\url{http://doi.org/10.1098/rstb.2006.1920}.

\hyperdef{}{ref-Oren2013}{\label{ref-Oren2013}}
4. \textbf{Oren A}, \textbf{Garrity GM}. 2013. Then and now: A
systematic review of the systematics of prokaryotes in the last
80~years. Antonie van Leeuwenhoek \textbf{106}:43--56.
doi:\url{http://doi.org/10.1007/s10482-013-0084-1}.

\hyperdef{}{ref-Brosius1978}{\label{ref-Brosius1978}}
5. \textbf{Brosius J}, \textbf{Palmer ML}, \textbf{Kennedy PJ},
\textbf{Noller HF}. 1978. Complete nucleotide sequence of a 16S
ribosomal RNA gene from \emph{Escherichia coli.} Proceedings of the
National Academy of Sciences \textbf{75}:4801--4805.
doi:\url{http://doi.org/10.1073/pnas.75.10.4801}.

\hyperdef{}{ref-Schloss2004}{\label{ref-Schloss2004}}
6. \textbf{Schloss PD}, \textbf{Handelsman J}. 2004. Status of the
microbial census. Microbiology and Molecular Biology Reviews
\textbf{68}:686--691.
doi:\url{http://doi.org/10.1128/mmbr.68.4.686-691.2004}.

\hyperdef{}{ref-Dykhuizen1998}{\label{ref-Dykhuizen1998}}
7. \textbf{Dykhuizen DE}. 1998. Santa Rosalia revisited: Why are there
so many species of bacteria? Antonie van Leeuwenhoek \textbf{73}:25--33.
doi:\url{http://doi.org/10.1023/a:1000665216662}.

\hyperdef{}{ref-Curtis2002}{\label{ref-Curtis2002}}
8. \textbf{Curtis TP}, \textbf{Sloan WT}, \textbf{Scannell JW}. 2002.
Estimating prokaryotic diversity and its limits. Proceedings of the
National Academy of Sciences \textbf{99}:10494--10499.
doi:\url{http://doi.org/10.1073/pnas.142680199}.

\hyperdef{}{ref-HMP2012}{\label{ref-HMP2012}}
9. \textbf{The Human Microbiome Consortium}. 2012. Structure, function
and diversity of the healthy human microbiome. Nature
\textbf{486}:207--214. doi:\url{http://doi.org/10.1038/nature11234}.

\hyperdef{}{ref-Gilbert2014}{\label{ref-Gilbert2014}}
10. \textbf{Gilbert JA}, \textbf{Jansson JK}, \textbf{Knight R}. 2014.
The Earth Microbiome Project: Successes and aspirations. BMC Biology
\textbf{12}:69. doi:\url{http://doi.org/10.1186/s12915-014-0069-1}.

\hyperdef{}{ref-AmaralZettler2010}{\label{ref-AmaralZettler2010}}
11. \textbf{Amaral-Zettler L}, \textbf{Artigas LF}, \textbf{Baross J},
\textbf{P.A. LB}, \textbf{Boetius A}, \textbf{Chandramohan D},
\textbf{Herndl G}, \textbf{Kogure K}, \textbf{Neal P},
\textbf{Pedrós-Alió C}, \textbf{Ramette A}, \textbf{Schouten S},
\textbf{Stal L}, \textbf{Thessen A}, \textbf{Leeuw J de}, \textbf{Sogin
M}. 2010. A global census of marine microbes, pp. 221--245. \emph{In}
Life in the worlds oceans. Wiley-Blackwell.

\hyperdef{}{ref-Youssef2009}{\label{ref-Youssef2009}}
12. \textbf{Youssef N}, \textbf{Sheik CS}, \textbf{Krumholz LR},
\textbf{Najar FZ}, \textbf{Roe BA}, \textbf{Elshahed MS}. 2009.
Comparison of species richness estimates obtained using nearly complete
fragments and simulated pyrosequencing-generated fragments in 16S rRNA
gene-based environmental surveys. Applied and Environmental Microbiology
\textbf{75}:5227--5236. doi:\url{http://doi.org/10.1128/aem.00592-09}.

\hyperdef{}{ref-Schloss2010}{\label{ref-Schloss2010}}
13. \textbf{Schloss PD}. 2010. The effects of alignment quality,
distance calculation method, sequence filtering, and region on the
analysis of 16S rRNA gene-based studies. PLoS Comput Biol
\textbf{6}:e1000844.
doi:\url{http://doi.org/10.1371/journal.pcbi.1000844}.

\hyperdef{}{ref-Alivisatos2015}{\label{ref-Alivisatos2015}}
14. \textbf{Alivisatos AP}, \textbf{Blaser MJ}, \textbf{Brodie EL},
\textbf{Chun M}, \textbf{Dangl JL}, \textbf{Donohue TJ},
\textbf{Dorrestein PC}, \textbf{Gilbert JA}, \textbf{Green JL},
\textbf{Jansson JK}, \textbf{Knight R}, \textbf{Maxon ME},
\textbf{McFall-Ngai MJ}, \textbf{Miller JF}, \textbf{Pollard KS},
\textbf{Ruby EG}, \textbf{Taha SA}. 2015. A unified initiative to
harness Earth's microbiomes. Science \textbf{350}:507--508.
doi:\url{http://doi.org/10.1126/science.aac8480}.

\hyperdef{}{ref-Liux5f2012}{\label{ref-Liux5f2012}}
15. \textbf{Li E}, \textbf{Hamm CM}, \textbf{Gulati AS}, \textbf{Sartor
RB}, \textbf{Chen H}, \textbf{Wu X}, \textbf{Zhang T}, \textbf{Rohlf
FJ}, \textbf{Zhu W}, \textbf{Gu C}, \textbf{Robertson CE}, \textbf{Pace
NR}, \textbf{Boedeker EC}, \textbf{Harpaz N}, \textbf{Yuan J},
\textbf{Weinstock GM}, \textbf{Sodergren E}, \textbf{Frank DN}. 2012.
Inflammatory bowel diseases phenotype, textitC. difficile and NOD2
genotype are associated with shifts in human ileum associated microbial
composition. PLoS ONE \textbf{7}:e26284.
doi:\url{http://doi.org/10.1371/journal.pone.0026284}.

\hyperdef{}{ref-Kongux5f2012}{\label{ref-Kongux5f2012}}
16. \textbf{Kong HH}, \textbf{Oh J}, \textbf{Deming C}, \textbf{Conlan
S}, \textbf{Grice EA}, \textbf{Beatson MA}, \textbf{Nomicos E},
\textbf{Polley EC}, \textbf{Komarow HD}, \textbf{Murray PR},
\textbf{Turner ML}, \textbf{Segre JA}. 2012. Temporal shifts in the skin
microbiome associated with disease flares and treatment in children with
atopic dermatitis. Genome Research \textbf{22}:850--859.
doi:\url{http://doi.org/10.1101/gr.131029.111}.

\hyperdef{}{ref-Griceux5f2009}{\label{ref-Griceux5f2009}}
17. \textbf{Grice EA}, \textbf{Kong HH}, \textbf{Conlan S},
\textbf{Deming CB}, \textbf{Davis J}, \textbf{Young AC},
\textbf{Bouffard GG}, \textbf{Blakesley RW}, \textbf{Murray PR},
\textbf{Green ED}, \textbf{Turner ML}, \textbf{Segre JA}. 2009.
Topographical and temporal diversity of the human skin microbiome.
Science \textbf{324}:1190--1192.
doi:\url{http://doi.org/10.1126/science.1171700}.

\hyperdef{}{ref-Griceux5f2010}{\label{ref-Griceux5f2010}}
18. \textbf{Grice EA}, \textbf{Snitkin ES}, \textbf{Yockey LJ},
\textbf{Bermudez DM}, \textbf{Liechty KW}, \textbf{Segre JA},
\textbf{Mullikin J}, \textbf{Blakesley R}, \textbf{Young A}, \textbf{Chu
G}, \textbf{Ramsahoye C}, \textbf{Lovett S}, \textbf{Han J},
\textbf{Legaspi R}, \textbf{Fuksenko T}, \textbf{Reddix-Dugue N},
\textbf{Sison C}, \textbf{Gregory M}, \textbf{Montemayor C},
\textbf{Gestole M}, \textbf{Hargrove A}, \textbf{Johnson T},
\textbf{Myrick J}, \textbf{Riebow N}, \textbf{Schmidt B},
\textbf{Novotny B}, \textbf{Gupti J}, \textbf{Benjamin B},
\textbf{Brooks S}, \textbf{Coleman H}, \textbf{Ho S-l},
\textbf{Schandler K}, \textbf{Smith L}, \textbf{Stantripop M},
\textbf{Maduro Q}, \textbf{Bouffard G}, \textbf{Dekhtyar M},
\textbf{Guan X}, \textbf{Masiello C}, \textbf{Maskeri B},
\textbf{McDowell J}, \textbf{Park M}, \textbf{Thomas PJ}. 2010.
Longitudinal shift in diabetic wound microbiota correlates with
prolonged skin defense response. Proceedings of the National Academy of
Sciences \textbf{107}:14799--14804.
doi:\url{http://doi.org/10.1073/pnas.1004204107}.

\hyperdef{}{ref-Kirkux5fHarrisux5f2012}{\label{ref-Kirkux5fHarrisux5f2012}}
19. \textbf{Harris JK}, \textbf{Caporaso JG}, \textbf{Walker JJ},
\textbf{Spear JR}, \textbf{Gold NJ}, \textbf{Robertson CE},
\textbf{Hugenholtz P}, \textbf{Goodrich J}, \textbf{McDonald D},
\textbf{Knights D}, \textbf{Marshall P}, \textbf{Tufo H}, \textbf{Knight
R}, \textbf{Pace NR}. 2012. Phylogenetic stratigraphy in the guerrero
negro hypersaline microbial mat. The ISME Journal \textbf{7}:50--60.
doi:\url{http://doi.org/10.1038/ismej.2012.79}.

\hyperdef{}{ref-Sogin2006}{\label{ref-Sogin2006}}
20. \textbf{Sogin ML}, \textbf{Morrison HG}, \textbf{Huber JA},
\textbf{Welch DM}, \textbf{Huse SM}, \textbf{Neal PR}, \textbf{Arrieta
JM}, \textbf{Herndl GJ}. 2006. Microbial diversity in the deep sea and
the underexplored ``rare biosphere''. Proceedings of the National
Academy of Sciences \textbf{103}:12115--12120.
doi:\url{http://doi.org/10.1073/pnas.0605127103}.

\hyperdef{}{ref-Whitman1998}{\label{ref-Whitman1998}}
21. \textbf{Whitman WB}, \textbf{Coleman DC}, \textbf{Wiebe WJ}. 1998.
Prokaryotes: The unseen majority. Proceedings of the National Academy of
Sciences \textbf{95}:6578--6583.
doi:\url{http://doi.org/10.1073/pnas.95.12.6578}.
